---
title: 'Flexible AFT models II: Bayesian + skew-symmetric'
author: "[F Javier Rubio](https://sites.google.com/site/fjavierrubio67/)"
date: "25/06/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bayesian AFT models with skew-symmetric errors

[Log-skew-symmetric distributions](http://onlinelibrary.wiley.com/doi/10.1002/env.1004/abstract) are univariate models with positive support obtained by transforming the class of [skew-symmetric distributions](http://www.jstor.org/stable/24307231). This class of distributions is very flexible, easy to implement, and contains members that can capture different tail behaviours and shapes, producing also a variety of hazard functions. These distributions represent a flexible alternative to the classical choices such as the log-normal, Gamma, and Weibull distributions. Below, we present an application using real data in the contexts of lung cancer survival using Bayesian [Accelerated Failure Time models](https://en.wikipedia.org/wiki/Accelerated_failure_time_model).:

**References**

1. [Bayesian linear regression with skew-symmetric error distributions with applications to survival analysis](http://dx.doi.org/10.1002/sim.6897)


```{r}
rm(list=ls())

# Required packages
library(survival)
library(Rtwalk)
library(sn)
library(TeachingDemos)

# Lung cancer data
data(lung)
str(lung)

vnames = c("age","sex","ph.ecog")

X = cbind(1,lung$age,lung$sex,lung$ph.ecog)[-14,]
logY = log(lung$time)[-14]
n = length(logY)

status = (1-(lung$status-1))[-14]
#######################################################################################################################################
# Prior on lambda
#######################################################################################################################################

c = pi/2
lpriorlambda = function(lambda) return( dt(c*lambda,df=0.5,log=T) + log(c) )

#######################################################################################################################################
# Log Skew Normal AFT Model
#######################################################################################################################################

# Log-likelihood
llsn <- function(par){
if(par[5]>0){
var1 = log(dsn(logY - X%*%par[1:4],0,par[5],par[6])^(1-status)) 
var2 = log( (1 - psn(logY - X%*%par[1:4],0,par[5],par[6]))^(status)  )
return(-sum(var1+var2))
}
return(Inf)
} 

# Maximum Likelihood Estimation
OPT1 = optim(c(7,0,0,0,1,-2),llsn,method="BFGS",control=list(maxit=10000))

init=OPT1$par

# Log Posterior

lpsl <- function(par){
var1 = log(dsn(logY - X%*%par[1:4],0,par[5],par[6])^(1-status)) 
var2 = log( (1 - psn(logY - X%*%par[1:4],0,par[5],par[6]))^(status)  )
#return(-sum(var1+var2)+log(par[5]) )
ifelse(is.na(sum(var1+var2)),
return(10000000),
return(-sum(var1+var2)+log(par[5]) - lpriorlambda(par[6]) ))
} 

# Simulation from the posterior

Support <- function(x) {
	((0.1 < x[5]) )	
}

X0 <- function(x) { init + 0.1*runif(6,-0.1,0.1) }

# It takes a couple of minutes to run
set.seed(1234)
info <- Runtwalk( dim=6,  Tr=105000,  Obj=lpsl, Supp=Support, x0=X0(), xp0=X0(), PlotLogPost = FALSE) 

# Thinning and burning the chain
ind=seq(5000,105000,50)
str(info)

beta1 = info$output[,1][ind]
beta2 = info$output[,2][ind]
beta3 = info$output[,3][ind]
beta4 = info$output[,4][ind]
sigma = info$output[,5][ind]
lambda = info$output[,6][ind]

# Some histograms of the posterior samples

hist(beta1)
hist(beta2)
hist(beta3)
hist(beta4)
hist(sigma)
hist(lambda)

#########################################################################
# Savage-Dickey ratio for log-skew-normal vs log-normal
#########################################################################

h = bw.nrd0(lambda)
postgammav <- Vectorize(function(x) mean(dnorm((x-lambda)/h)/h))
curve(postgammav,-10,1)

BF1 = postgammav(0)/exp(lpriorlambda(0))

# Bayes factor against the model with normal (instead of skew-normal) errors
# Strong evidence in favour of the log-skew-normal model
BF1
log(BF1)
```
