---
title: "Flexible AFT models I: MLE"
author: "[F Javier Rubio](https://sites.google.com/site/fjavierrubio67/)"
date: "25/06/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Accelerated Failure Time models with flexible errors

Log-twopiece distributions are univariate models with positive support obtained by transforming the class of [two-piece distributions](https://rpubs.com/FJRubio/twopiece). This class of distributions is very flexible, easy to implement, and contains members that can capture different tail behaviours and shapes, producing also a variety of hazard functions. These distributions represent a flexible alternative to the classical choices such as the log-normal, Gamma, and Weibull distributions. Below, we present an application using real data in the contexts of lung cancer survival using [Accelerated Failure Time models](https://en.wikipedia.org/wiki/Accelerated_failure_time_model).

**References.**

1. [Inference in Two-Piece Location-Scale Models with Jeffreys Priors](http://projecteuclid.org/euclid.ba/1393251764)

2. [Survival and lifetime data analysis with a flexible class of distributions](http://dx.doi.org/10.1080/02664763.2015.1120710)

```{r}
rm(list=ls())

# Required packages
library(survival)
library(twopiece)

# Lung Cancer data
data(lung)
str(lung)

vnames = c("age","sex","ph.ecog")

# Remove a missing observation
X = cbind(1,lung$age,lung$sex,lung$ph.ecog)[-14,]
logY = log(lung$time)[-14]
n = length(logY)

status = (1-(lung$status-1))[-14]

#######################################################################################################################################
# Weibull, lognormal and log-logistic fit using survreg
#######################################################################################################################################

summary( survreg(Surv(time, status) ~ age+sex+ph.ecog, data=cancer, dist="weibull") )

summary( survreg(Surv(time, status) ~ age+sex+ph.ecog, data=cancer, dist="lognormal") )

summary( survreg(Surv(time, status) ~ age+sex+ph.ecog, data=cancer, dist="loglogistic") )

#######################################################################################################################################
# Log Normal fit
#######################################################################################################################################

lln <- function(par){
if(par[5]>0){
var1 = log(dnorm(logY - X%*%par[1:4],0,par[5])^(1-status))
var2 = log( (1 - pnorm(logY - X%*%par[1:4],0,par[5]))^(status)  )
return(-sum(var1+var2))
}
return(Inf)
} 

OPT4 = optim(c(6,0,0,0,1),lln,control=list(maxit=10000))
AIC4 = 2*OPT4$value + 2*5

# MLE
OPT4$par
#######################################################################################################################################
# Log two-piece normal fit
#######################################################################################################################################

lltpn <- function(par){
if(par[5]>0 & par[6]>-1 & par[6]<1){
var1 = log(dtp3(logY - X%*%par[1:4],0,par[5],par[6],param="eps",FUN=dnorm)^(1-status) )
var2 = log( (1 - ptp3(logY - X%*%par[1:4],0,par[5],par[6],param="eps",FUN=pnorm,log.p=F) )^(status) )
return(-sum(var1+var2))
}
return(Inf)
} 

OPT2 = optim(c(OPT4$par,0.3),lltpn,control=list(maxit=10000))
AIC2 = 2*OPT2$value + 2*6

# MLE
OPT2$par

#######################################################################################################################################
# Log Logistic fit
#######################################################################################################################################

lll <- function(par){
if(par[5]>0){
var1 = log(dlogis(logY - X%*%par[1:4],0,par[5])^(1-status))
var2 = log( (1 - plogis(logY - X%*%par[1:4],0,par[5]))^(status)  )
return(-sum(var1+var2))
}
return(Inf)
} 

OPT5 = optim(c(6,0,0,0,0.5),lll,control=list(maxit=10000))
AIC5 = 2*OPT5$value + 2*5

# MLE
OPT5$par

#######################################################################################################################################
# Log two-piece logistic fit 
#######################################################################################################################################

lltpl <- function(par){
if(par[5]>0 & par[6]>-1 & par[6]<1){
var1 = log(dtp3(logY - X%*%par[1:4],0,par[5],par[6],param="eps",FUN=dlogis)^(1-status)) 
var2 = log( (1 - ptp3(logY - X%*%par[1:4],0,par[5],par[6],param="eps",FUN=plogis,log.p=F))^(status)  )
return(-sum(var1+var2))
}
return(Inf)
} 

OPT6 = optim(c(OPT5$par,0.2),lltpl,control=list(maxit=10000))
AIC6 = 2*OPT6$value + 2*6

# MLE
OPT6$par

# Comparison using AIC

c(AIC2,AIC4,AIC5,AIC6)
```
