---
title: 'Flexible AFT Models III: Bayesian + two-piece'
author: '[F. Javier Rubio](https://sites.google.com/site/fjavierrubio67)'
date: '25/06/2017'
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bayesian AFT models with two-piece errors

[Log-two piece distributions](http://onlinelibrary.wiley.com/doi/10.1002/env.1004/abstract) are univariate models with positive support obtained by transforming the class of [two-piece distributions](https://rpubs.com/FJRubio/twopiece) [@rubio:2020]. This class of distributions is very flexible, easy to implement, and contains members that can capture different tail behaviours and shapes, producing also a variety of hazard functions. These distributions represent a flexible alternative to the classical choices such as the log-normal, Gamma, and Weibull distributions. Below, we present an application using real data in the contexts of small-cell lung cancer survival using Bayesian [Accelerated Failure Time models](https://en.wikipedia.org/wiki/Accelerated_failure_time_model). Linear regression models with two-piece errors are closely related to [quantile regression](https://en.wikipedia.org/wiki/Quantile_regression) (for two-piece Laplace errors) and [asymmetric least squares](http://www.jstor.org/stable/1911031) (for two-piece normal errors): see [Tractable Bayesian variable selection: beyond normality](https://doi.org/10.1080/01621459.2017.1371025) for more details.

# References

1. [Flexible objective Bayesian linear regression with applications in survival analysis](https://doi.org/10.1080/02664763.2016.1182138)

2. [Survival and lifetime data analysis with a flexible class of distributions](https://doi.org/10.1080/02664763.2015.1120710)

# R code
```{r}
rm(list=ls())

library(survival)
library(Rtwalk)
library(twopiece)
library(TeachingDemos)
library(LaplacesDemon)
library(emplik)



# Small Cell Cancer Data

data(smallcell)
str(smallcell)

vnames = c("age","arm")

X = cbind(1,smallcell$entry,smallcell$arm)
logY = log(smallcell$survival)
n = length(logY)

status = 1-smallcell$indicator

#######################################################################################################################################
# Logistic fit
#######################################################################################################################################

summary( survreg(Surv(survival,smallcell$indicator) ~ entry+arm, data=smallcell, dist="loglogistic") )

#######################################################################################################################################
# Log Skew Laplace AFT Model
#######################################################################################################################################

plaplace1 <- function(q, location = 0, scale = 1,log.p=FALSE) plaplace(q, location = 0, scale = 1)

# Log likelihood
llsl <- function(par){
if(par[4]>0 & par[5]<1 & par[5]>-1){
var1 = log(dtp3(logY - X%*%par[1:3],0,par[4],par[5],param="eps",FUN=dlaplace)^(1-status)) 
var2 = log( (1 - ptp3(logY - X%*%par[1:3],0,par[4],par[5],param="eps",FUN=plaplace1))^(status)  )
return(-sum(var1+var2))
}
return(Inf)
} 


# Maximum Likelihood Estimation

OPT1 = optim(c(7,0,0,1,0),llsl,method="BFGS",control=list(maxit=10000))

init=OPT1$par

# Log Posterior

lpsl <- function(par){
var1 = log(dtp3(logY - X%*%par[1:3],0,par[4],par[5],param="eps",FUN=dlaplace)^(1-status)) 
var2 = log( (1 - ptp3(logY - X%*%par[1:3],0,par[4],par[5],param="eps",FUN=plaplace1))^(status)  )
return(-sum(var1+var2)+log(par[4])  )
} 

# Simulation from the posterior


Support <- function(x) {
	((0.1 < x[4])&(-1<x[5])&(x[5]<1))	
}

X0 <- function(x) { init + 0.1*runif(5,-0.1,0.1) }

set.seed(1234)
info <- Runtwalk( dim=5,  Tr=105000,  Obj=lpsl, Supp=Support, x0=X0(), xp0=X0(),PlotLogPost = FALSE) 

# Thinning and burning the chain

ind=seq(5000,105000,25)
str(info)


# Some histograms of the posterior samples

beta1 = info$output[,1][ind]
beta2 = info$output[,2][ind]
beta3 = info$output[,3][ind]
sigma = info$output[,4][ind]
gamma = info$output[,5][ind]


hist(beta1)
hist(beta2)
hist(beta3)
hist(sigma)
hist(gamma)


#########################################################################
# Savage-Dickey ratio for log-skew/twopiece-Laplace vs log-Laplace
#########################################################################

h = bw.nrd0(gamma)
postgammav <- Vectorize(function(x) mean(dnorm((x-gamma)/h)/h))
curve(postgammav,-1,1)

BF1 = postgammav(0)/dunif(0,-1,0)

# Bayes factor against the model with Laplace (instead of skew/twopiece-Laplace) errors
# Strong evidence in favour of the log-skew-Laplace model
BF1
log(BF1)
```
